---
output: pdf_document
---

##Title: Measuring Impact of Natural Events 
##Snopysis 
This analysis measures the impact of natural events.  The two metrics used to measure the impact of the natural events are fatalities and economic damage.  

##Data Processing
```{r echo=TRUE}
require("data.table")
setwd("C:/Users/Eric.Kim/Desktop/TSA Files/Coursera/Reproducible Research/Assignment 2")

#read bz2 file
storm<-read.csv("repdata-data-StormData.csv.bz2")

#aggregate fatalities by event
storm<-data.table(storm)
setkey(storm, EVTYPE)
health<-data.frame(storm[,sum(FATALITIES), by=EVTYPE])
names(health)[1]<-"event"
names(health)[2]<-"fatalities"
health<-health[with(health, order(-fatalities)),]
health<-head(health,20)
```

```{r}
#convert storm from data table to data frame
storm<-data.frame(storm)

#Transform the costs data by multipling by thousands or millions.  The reason for the transformation 
#is to add the crop and property damage as total costs
storm$CROPDMG2<-ifelse(storm$CROPDMGEXP=='K', storm$CROPDMG*1000, 
       ifelse(storm$CROPDMGEXP=='M', storm$CROPDMG*1000000, storm$CROPDMG))
storm$PROPDMG2<-ifelse(storm$PROPDMGEXP=='K', storm$PROPDMG*1000, 
       ifelse(storm$PROPDMGEXP=='M', storm$PROPDMG*1000000, storm$PROPDMG))

#sum damage costs
storm$damage<-rowSums(storm[,38:39])
storm<-data.table(storm)
setkey(storm, EVTYPE)
damage<-data.frame(storm[,sum(damage), by=EVTYPE])
names(damage)[1]<-"event"
names(damage)[2]<-"damage"
damage<-damage[with(damage, order(-damage)),]
damage<-head(damage,20)
```

##Results
The most destructive natural events are tornados.  They caused the most fatalities and economic damage is Tornado, resulting in 5,633 deaths and causing $52B in damages.  

###Events Most Harmful to Public Health
```{r}
#display to 20 fatal events
health  
```

```{r}
#plot fatalities by event 
library(ggplot2)
ggplot(health, aes(event, fatalities))+geom_bar(stat="identity")+theme(axis.text.x = element_text(angle = 90, hjust = 1))+ggtitle("Events Most Harmful to Public Health")
``` 
 
Figure Description: The figure above lists the top 20 events that have caused the most fatalities

###Events with the Greatest Economic Damage
```{r}
#display events that cause the most economic damaging 
damage
```


```{r}
#graph economic damage
ggplot(damage, aes(event, damage))+geom_bar(stat="identity")+theme(axis.text.x = element_text(angle = 90, hjust = 1))+ggtitle("Events that Cause the Most Economic Damage")
``` 
 
Figure Description: The figure above lists the top 20 events that have caused the most economic damage